\ProjectName{}: Data Imputation as a Query Optimization

### 2017-02-28 paper/experiments.tex
comments

@@

\begin{figure}
\includegraphics[width=\columnwidth]{figures/running_times_combined.png}
\caption{Query runtimes with imputation on base tables, \ProjectName{}
    optimizing for quality ($\alpha=0$), and \ProjectName{} optimizing for
    performance ($\alpha=1$). Quality-optimized, on-the-fly imputation provides an order of
    magnitude speedup over imputing on base tables. Efficiency-optimized on-the-fly
    imputation can provide another order-of-magnitude speedup at the expense of quality. For
    relatively expensive queries over large datasets (Query 9), runtime of imputation can be reduced from 6.5
    hours to 7 seconds. }
\label{fig:runtimes}
\end{figure}

@@

\Cref{table:smape} shows the Symmetric-Mean-Absolute-Percentage-Error (SMAPE) for \ProjectName{}'s query results when compared to running imputation on the base tables and executing the query on the cleaned data \todo{SMAPE citation}.
Each query was separately run 200 times in both settings and then results from the two approaches were paired up and compared tuple-wise.
We average tuple-wise absolute percentage deviations within each iteration of a query, and we report this value averaged over all iterations.
We can see that optimizing for quality indeed reduces the SMAPE of query results.
In general, the SMAPE relative to the base-imputation approach are low in all cases\todo{give actual value}, indicating that on-the-fly imputation produces similar results to imputation at the base tables.

@@

In many real-world cases, applying the imputation step at the base table is prohibitively
expensive.
To illustrate the increasing difficulty of such an approach as datasets scale, we run~\Cref{q9} over the ACS dataset.
Applying the imputation operation to the base table is extremely expensive, in our case
completing in 355 minutes(\Cref{fig:runtimes})\todo{Micah: this doesn't match the 6.5 hours in the figure}. In contrast, \ProjectName{} executes a quality-optimized version
in 7 seconds and a runtime-optimized version in 1 second. This highlights the potential
benefit of using our system for early data exploration.

>>>>

\subsection{Results}

@@

\begin{figure}
\includegraphics[width=\columnwidth]{figures/running_times_combined.png}
\caption{Query runtimes with imputation on base tables, \ProjectName{}
    optimizing for quality ($\alpha=0$), and \ProjectName{} optimizing for
    performance ($\alpha=1$). Quality-optimized, on-the-fly imputation provides an order of
    magnitude speedup over imputing on base tables. Efficiency-optimized on-the-fly
    imputation can provide another order-of-magnitude speedup at the expense of quality. For
    relatively expensive queries over large datasets (Query 9), runtime of imputation can be reduced from 6.5
    hours to 7 seconds. \srm{make this a bar graph}}
\label{fig:runtimes}
\end{figure}

@@

\subsubsection{Accuracy vs. Base-table Imputation}

\Cref{table:smape} shows the Symmetric-Mean-Absolute-Percentage-Error (SMAPE) for \ProjectName{}'s query results when compared to running imputation on the base tables and executing the query on this cleaned base data \todo{SMAPE citation}.
Each query was separately run 200 times in both settings and then results from the two approaches were paired up and compared tuple-wise.
We average tuple-wise absolute percentage deviations within each iteration of a query, and we report this value averaged over all iterations.
We can see that optimizing for quality indeed reduces the SMAPE of query results.
In general, the SMAPE relative to the base-imputation approach are low in all cases\todo{give actual value}, indicating that on-the-fly imputation produces similar results to imputation at the base tables.

@@

\subsubsection{Runtime of Base-table Imputation}
In many real-world cases, applying the imputation step at the base table is prohibitively
expensive.
To illustrate the increasing difficulty of such an approach as datasets scale, we run~\Cref{q9} over the ACS dataset.
Applying the imputation operation to the base table is extremely expensive, in our case
completing in 355 minutes(\Cref{fig:runtimes})\todo{Micah: this doesn't match the 6.5 hours in the figure}. In contrast, \ProjectName{} executes a quality-optimized version
in 7 seconds and a runtime-optimized version in 1 second. This highlights the potential
benefit of using our system for early data exploration.

\srm{note somewhere that imputation requires both training and running model, and state why this is reasonable / the only choice}

### 2017-02-28 paper/introduction.tex
comments

% Solution
One common approach is to manually replace missing values using some sort of statistical or predictive model based on other values in the table and record.
This process is called \emph{imputation}.
In this paper, we present a new system, called \ProjectName{}, which is the designed to selectively apply imputation to a subset of records, on-the-fly, during query execution.  \emph{The key insight behind \ProjectName{} is that imputation
only needs to be performed on the data relevant to a particular query and
that this subset is generally much smaller than the entire database.}  
% Background
While traditional imputation methods work over the entire data set and replace all missing values, running a sophisticated imputation algorithm over a large data set can be very expensive -- out experiments show that even a relatively simple regression algorithm algorithms tables about 75 minutes to train and run
on a 600K row database;  prior work~\cite{AkandeEtAl} reports times of 45 minutes to impute missing values on a 1\% sample of an XXX row database.
A simpler approach might drop all rows with missing values, but not only can this introduce bias into the results, but this may result in discarding all of the data.
In contrast to existing systems~\cite{burgette2010multiple,akande2015empirical}, \ProjectName{} avoids imputing over the entire data set, as done with traditional methods. Specifically, \ProjectName{}
carefully plans the placement  imputation / row drop operations inside the query plan, resulting in a 
significant speedup in query execution while minimizing the number of dropped rows.  Also unlike previous work, the focus on of our work is not on the imputation algorithms themselves (we can employ almost any such algorithm), but rather on the  query planning involved in placement of imputation operations
in an optimal way into query plans.   Specifically, our optimization algorithms generate the Pareto-optimal frontier of plans that trade imputation cost for result quality.

>>>>

% Solution
One common approach is to manually replace missing values using some sort of statistical or predictive model based on other values in the table and record.
This process is called \emph{imputation}.
In this paper, we present a new system, called \ProjectName{}, which is the designed to selectively apply imputation to a subset of records, on-the-fly, during query execution.  \emph{The key insight behind \ProjectName{} is that imputation
only needs to be performed on the data relevant to a particular query and
that this subset is generally much smaller than the entire database.}  
% Background
While traditional imputation methods work over the entire data set and replace all missing values, running a sophisticated imputation algorithm over a large data set can be very expensive -- out experiments show that even a relatively simple regression algorithm algorithms tables about XXX minutes to train and run
on a 600K row database;  prior work~\cite{AkandeEtAl} reports times of 45 minutes to impute missing values on a 1\% sample of an XXX row database.
\srm{Make numbers above consistent with experiments.}
A simpler approach might drop all rows with missing values, but not only can this introduce bias into the results, but this may result in discarding all of the data.
In contrast to existing systems~\cite{burgette2010multiple,akande2015empirical}, \ProjectName{} avoids imputing over the entire data set, as done with traditional methods. Specifically, \ProjectName{}
carefully plans the placement  imputation / row drop operations inside the query plan, resulting in a 
significant speedup in query execution while minimizing the number of dropped rows.  Also unlike previous work, the focus on of our work is not on the imputation algorithms themselves (we can employ almost any such algorithm), but rather on the  query planning involved in placement of imputation operations
in an optimal way into query plans.   Specifically, our optimization algorithms generate the Pareto-optimal frontier of plans that trade imputation cost for result quality.
